Introducing Distributed Dynamic Data-intensive (D3) Science: Understanding Applications and Infrastructure

### 2016-08-17 3dpas_intro.tex
minor: removing claim about everything is broken, but this commit message..
Discussion about "hierarchy" and "granularity" of applications.

Starting with the experience of the attendees of the 3DPAS workshops,
we have assembled a set of D3 ``applications'' that we describe in
this section.  We realize that we have overloaded the term application
here.  In this document, some of our applications are really {\em
  traditional applications}, meaning a program that is developed by a
user or project to try to answer a science question, usually with some
input data needed and some output data produced.  Applications in this
category are: NGS (next generation sequencing) Analytics (\S\ref{bioSilvia}), CMB (cosmic microwave background)
(\S\ref{astroJulian}), Fusion (\S\ref{fusionScott}), Industrial
Incident Notification and Response (\S\ref{Kees}), MODIS (moderate resolution imaging spectroradiometer) Data
Processing (\S\ref{modisKeith}), and Distributed Network Intrusion
Detection (\S\ref{detectionJon}).

@@

All algorithms, architectures, storage, and networks have broken or
will break soon.  Some NGS analytics are not good fits for core-rich
and memory-poor modern architectures. Some problems (e.g., comparing
k-mers across metagenomic data sets) need terabyte-scale memory, not
thousands of cores.~\cite{ngs-gap}

>>>>

Starting with the experience of the attendees of the 3DPAS workshops, we have
assembled a set of D3 ``applications'' that we describe in this section.  
%We realize that we have overloaded the term ``application'' here.

The term ``application'' is often overloaded in literature and the computational
science discourse.  Sometimes the term is a reference to a stand-alone,
independent executable (e.g., a Molecular Dynamics simulation kernel such as
AMBER or Gromacs). Sometimes an application refers to science problem to which
the executable is put to use (e.g., compute the free-energy of binding energy of
a drug candidate), but sometimes it is also used to reference the end-to-end
workflow (i.e., a multi-stage computational process requiring a sequence of
possibly distinct simulations and executables).  To add chaos to the confusion,
applications are sometimes also as a proxy for projects (e.g., the ATLAS
project), where the project could in of itself possibly be a series of
independent applications, each with application in turn a reference to a defined
science objective achieved using a set of executables.

In addition to the aforementioned hierarchy and granularity to which the term
application has been applied, the class of applications in any level differ
widely in scale, sophistication and type as well different objective and
objects.  Thus, it is inevitable that the non-contextual use of the word
application be a recipe for befuddlement.

In this document, some of our applications are really {\em traditional
  applications}, meaning a program that is developed by a user or project to try
to answer a science question, usually with some input data needed and some
output data produced.  Applications in this category are: NGS (next generation
sequencing) Analytics (\S\ref{bioSilvia}), CMB (cosmic microwave background)
(\S\ref{astroJulian}), Fusion (\S\ref{fusionScott}), Industrial Incident
Notification and Response (\S\ref{Kees}), MODIS (moderate resolution imaging
spectroradiometer) Data Processing (\S\ref{modisKeith}), and Distributed Network
Intrusion Detection (\S\ref{detectionJon}).

@@

% All algorithms, architectures, storage, and networks have broken or
% will break soon. 
Some NGS analytics are not good fits for core-rich and memory-poor modern
architectures. Some problems (e.g., comparing k-mers across metagenomic data
sets) need terabyte-scale memory, not thousands of cores.~\cite{ngs-gap}
